Tropical Andes Functional diversity for plants and Frugivores

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
working on plotting - not working still, update input files and object names

```{r}
# objects for functional diversity
PAM_plant_site_final <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
PAM_frugivore_site_final <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
plant_traits_df_final <- readRDS(file = file.path(data_path_L1,"plant_traits_df_final.rds"))
frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
site_loc_key_plant <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant.rds"))
site_loc_key_frugivore <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
```

@@

## Summary of the assemblages * species dataframe
```{r}
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
```

@@

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
```

@@

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_frugivore <- alpha_fd_indices_frugivore$"details"

plots_alpha_frugivore <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

 #comparing convex hulls of different hulls
plots_alpha_frugivore$"fric"$"patchwork"

plots_alpha_frugivore$"fdiv"$"patchwork"

plots_alpha_frugivore$"fspe"$"patchwork"

plots_alpha_frugivore$"fdis"$"patchwork"

plots_alpha_frugivore$"fori"$"patchwork"
```

@@

@@

@@

@@

>>>>

```{r}
plant_traits_df_final <- readRDS(file = file.path(data_path_L1,"plant_traits_df_final.rds"))
frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
```

```{r}
site_loc_key_plant_100km <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant_100km.rds"))
site_loc_key_frugivore_100km <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore_100km.rds"))
PAM_plant_site_final_100km <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final_100km.rds"))
PAM_frugivore_site_final_100km <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final_100km.rds"))
```

@@

## Summary of the assemblages * species dataframe
```{r}
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final_100km)
```

@@

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_frugivore <- rowSums(PAM_frugivore_site_final_100km)
subset_matrix_frugivore <- PAM_frugivore_site_final_100km[row_sums_frugivore >= 4, ]
```

@@

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = "fdis",
  #did not run fdiv or fric (slows it down); c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe")
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_frugivore <- alpha_fd_indices_frugivore$"details"
```

@@

```{r}
# Generate coordinates
subset_coords_frugivore <-site_loc_key_frugivore_100km[rowSums(PAM_frugivore_site_final_100km) >= 4, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

@@

```{r}
# Convert the dataframe to sf format
spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(5389)
```

@@

# Aggregate functional dispersion within each grid cell
aggregate_to_grid <- function(grid, points, value_col) {
  points %>%
    st_join(grid) %>%
    group_by(cellid) %>%
    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
    ungroup() %>%
    st_set_geometry("geometry")
}

# use funtion
resolution_meters <- 100000
variable_name <- "fdis_frugivore"

@@

# Convert geometries to centroids
aggregated_data_centroids <- st_centroid(aggregated_data)

# Extract coordinates from centroids
aggregated_data_coords <- st_coordinates(aggregated_data_centroids)

# Add coordinates as separate columns
aggregated_data <- aggregated_data %>%
  mutate(x = aggregated_data_coords[, 1], y = aggregated_data_coords[, 2])

# Remove rows with NA values in mean_value
aggregated_data <- aggregated_data %>%
  filter(!is.na(mean_value))
```

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = grid, fill = "grey")+
    geom_sf(data = aggregated_data, aes(fill = mean_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue")) +
    theme(text = element_text(size = 12)) # Explicitly setting text size
FDisplot
```
